On the specific recursive structure of iterated parametricity

### 2025-10-26 paper/paper.tex
paper: strip extradeps(cohs)

\newcommandx{\deps}[2][1,2]{\prim{deps}[][#1][#2][][]}
\newcommandx{\extradeps}[2][1,2]{\prim{deps}[][#1][#2][][]}
\newcommandx{\depscohs}[2][1,2]{\prim{depscohs}[][#1][#2][][]}
\newcommandx{\extradepscohs}[2][1,2]{\prim{depscohs}[][#1][#2][][]}
\newcommandx{\depstype}[2][1,2]{\prim{depstype}[][#1][#2][][]}
\newcommandx{\extradepstype}[2][1,2]{\prim{depstype}[][#1][#2][][]}
\newcommandx{\depscohstype}[2][1,2]{\prim{depscohstype}[][#1][#2][][]}
\newcommandx{\extradepscohstype}[2][1,2]{\prim{depscohstype}[][#1][#2][][]}

@@

Formally, this can be described as:
\begin{equation*}
  \fontsize{7.2}{9}\selectfont
  \begin{array}{llcl}
    \framep[n][0,...,p]           &
    \multicolumn{3}{l}{
      \left(\begin{array}{l}
                \framep[n][0,...,p-1]:\frametype[n][0,...,p-1]                             \\
                \painting[n][0,...,p-1]:\paintingtype[n][0,...,p-1](\framep[n][0,...,p-1]) \\
                \restrf[n][0,...,p-1]:\restrftype[n][0,...,p-1]
                \left(\begin{array}{l}
                    \framep[n][0,...,p-1]   \\
                    \painting[n][0,...,p-1] \\
                  \end{array}\right)                                              \\
              \end{array}\right) : \frametype[n][0,...,p]
    }                                                                  \\
    \framep[n][0]                 &
    \left(\begin{array}{l}
              \unitpoint \\
              \unitpoint \\
              \unitpoint \\
            \end{array}\right)        & \defeq & (\unitpoint,unit)       \\
    \framep[n][0,...,p+1]         &
    \left(\begin{array}{l}
              \framep[n][0,...,p]   \\
              \painting[n][0,...,p] \\
              \restrf[n][0,...,p]   \\
            \end{array}\right)      & \defeq &
    \left(\begin{array}{l}
              \framep[n+1][0,...,p]
              \left(\begin{array}{l}
                  \framep[n+1][0,...,p-1]   \\
                  \painting[n+1][0,...,p-1] \\
                  \restrf[n+1][0,...,p-1]   \\
                \end{array}\right), \\
              \Sigma d:\framep[n+1][0,...,p]
              \left(\begin{array}{l}
                  \framep[n+1][0,...,p-1]   \\
                  \painting[n+1][0,...,p-1] \\
                  \restrf[n+1][0,...,p-1]   \\
                \end{array}\right).\, \\
              \quad \layer[n][p]
              \left(\begin{array}{l}
                  \framep[n+1][0,...,p-1]   \\
                  \painting[n+1][0,...,p-1] \\
                  \restrf[n+1][0,...,p-1]   \\
                \end{array}\right) \\
            \end{array}\right)                   \\
    \restrftype[n][0,...,p-1]     &
    \multicolumn{3}{l}{
      \left(\begin{array}{l}
                \framep[n][0,...,p-1]:\frametype[n][0,...,p-1]                             \\
                \painting[n][0,...,p-1]:\paintingtype[n][0,...,p-1](\framep[n][0,...,p-1]) \\
              \end{array}\right) : Type_{l}
    }                                                                  \\
    \restrftype[n][0,...,0-1]     &
    \left(\begin{array}{l}
              \unitpoint \\
              \unitpoint \\
            \end{array}\right)        & \defeq &
    \unittype                                                          \\
    \restrftype[n][0,...,(p+1)-1] &
    \left(\begin{array}{l}
              \framep[n][0,...,p]   \\
              \painting[n][0,...,p] \\
            \end{array}\right)      & \defeq &
    \begin{array}{l}
      \Sigma \restrf[n+1][0,...,p-1]
      :\restrftype[n+1][0,...,p-1]
      \left(\begin{array}{l}
                \framep[n+1][0,...,p-1]   \\
                \painting[n+1][0,...,p-1] \\
              \end{array}\right)                         .  \\
      \Pi q\leq n.\,\Pi \omega.\,\framep[n+1][p]
      \left(\begin{array}{l}
                \framep[n+1][0,...,p-1]   \\
                \painting[n+1][0,...,p-1] \\
                \restrf[n+1][0,...,p-1]   \\
              \end{array}\right) \rightarrow  \framep[n][p] \\
    \end{array} \\
    \layer[n][p]                  &
    \left(\begin{array}{l}
              \framep[n][0,...,p]     \\
              \painting[n][0,...,p]   \\
              \restrf[n+1][0,...,p-1] \\
            \end{array}\right)    & \defeq &
    \Pi\omega.\,\painting[n][p](\restrf[n][p](d))
  \end{array}
\end{equation*}

@@

At this stage, it is also convenient to use an abbreviation to combine the extra parts of $\framep$, a $\painting$ and a $\restrf$ in one block, which we call $\extradeps[n][p,...,n-1]$:

\begin{align*}
  \extradeps[n][p,..,n-1] (\deps[n][0,...,p-1]) \defeq
  \left(\begin{array}{l}
            \framep[n][p,...,n-1] :\frametype[n][p,...,n-1] \\
            \painting[n][p,...,n-1] :\paintingtype[n][p,...,n-1]
            (\framep[n][p,...,n-1])                         \\
            \restrf[n][p,...,n-1]  :\restrftype[n][p,...,n-1]
            \left(\begin{array}{l}
                \deps[n-1][0,...,p]         \\
                \framep[n-1][p+1,...,n-1]   \\
                \painting[n-1][p+1,...,n-1] \\
              \end{array}\right)               \\
          \end{array}\right)
\end{align*}

This allows to define $\painting[n][p]$ formally:
\begin{equation*}
  \fontsize{7.2}{9}\selectfont
  \begin{array}{llcl}
    \painting[n][p]               &
    \multicolumn{3}{l}{
      \left(\begin{array}{l}
                \deps[n-1][0,...,p-1]:\depstype[n-1][0,...,p-1]           \\
                \extradeps[n-1][p,...,n-1]:\extradepstype[n-1][p,...,n-1] \\
                E:\framep[n-1][n-1]\rightarrow Type_l                     \\
              \end{array}\right) : \framep[n-1][n-1]\rightarrow Type_l
    }                                                                            \\
    \painting[n][n]               &
    \left(\begin{array}{l}
              \deps[n-1][0,...,n-1] \\
              \unitpoint            \\
              E                     \\
            \end{array}\right)      & \defeq & E                                   \\
    \painting[n][p](p<n)          &
    \left(\begin{array}{l}
              \deps[n-1][0,...,p-1]      \\
              \extradeps[n-1][p,...,n-1] \\
              E                          \\
            \end{array}\right) & \defeq & d \mapsto
    \left(\begin{array}{l}
              \Sigma l:(\Pi q\leq n.\,\Pi \omega.\
              \deps.\painting[n-1][0,...,p](\extradeps.\restrf[n][p](d))). \\
              \painting[n][p+1,...,n-1]
              \left(\begin{array}{l}
                  \deps[n-1][0,...,p]          \\
                  \extradeps[n-1][p+1,...,n-1] \\
                  E                            \\
                \end{array}\right)(d,l)                           \\
            \end{array}\right) \\
  \end{array}
\end{equation*}

@@

                \deps[n][0,...,p-2]:\depstype[n][0,...,p-2]                                \\
                \extradeps[n][p-1,...,n]:\extradepstype[n][p-1,...,n](\deps[n][0,...,p-1]) \\

@@

                    \deps[n][0,...,p-1]      \\
                    \extradeps[n][p-1,...,n] \\

@@

              \deps[n][0,...,p]        \\
              \extradeps[n][p-1,...,n] \\

@@

                \deps[n+1][0,...,p-1]    \\
                \extradeps[n][p-1,...,n] \\

@@

              \end{array}\right)     ,                                            \\

@@

                \deps[n][0,...,p-1]:\depstype[n][0,...,p-1]                                \\
                \extradeps[n][p-1,...,n]:\extradepstype[n][p-1,...,n](\deps[n][0,...,p-1]) \\

@@

              \end{array}\right)                         .     \\

@@

              \deps[n][0,...,p]:\depstype[n][0,...,p]                                    \\
              \extradeps[n][p-1,...,n]:\extradepstype[n][p-1,...,n](\deps[n][0,...,p-1]) \\

@@

At this new stage, it is convenient to use an abbreviation to combine sequences $\deps[n][0,..,p-1]$, $\extradeps[n][p,...,n]$, $\restrp[n][0,...,p-1]$ and $\cohf[n][0,...,p-1]$ in one block, which we call $\depscohs[n][0,...p-1]$:

\begin{align*}
  \depscohs[n][0,..,p-1] \defeq
  \left(\begin{array}{l}
            \deps[n][0,...,p-1] :\depstype[n][0,...,p-1]    \\
            \extradeps[n][0,...,p-1] :\extradepstype[n][p-1,...,n]
            (\deps[n][0,...,p-1])                           \\
            \restrp[n][0,...p-1]: \restrptype[n][0,...,p-1] \\
            \cohf[n][0,...,p-1]: \cohftype[n][0,...,p-1]    \\
          \end{array}\right)
\end{align*}
so that $\restrf[n][0,...,p]$ has type $\depscohs[n][0,..,p-1] \rightarrow \restrftype[n][0,...,p]$. Later, we shall write $\depscohs.\deps$, $\depscohs.\extradeps$, $\depscohs.\restrp$, and $\depscohs.\cohf$ for the corresponding projections of a $\depscohs$.

@@

                \depscohs[n][0,...,p-2]:\depstype[n][0,...,p-2]                                \\
                \extradepscohs[n][p-1,...,n]:\extradepstype[n][p-1,...,n](\deps[n][0,...,p-1]) \\

@@

                    \depscohs[n][0,...,p-1]      \\
                    \extradepscohs[n][p-1,...,n] \\

@@

              \depscohs[n][0,...,p]        \\
              \extradepscohs[n][p-1,...,n] \\

@@

                \depscohs[n+1][0,...,p-1]    \\
                \extradepscohs[n][p-1,...,n] \\

@@

                \depscohs[n][0,...,p-1]:\depscohstype[n][0,...,p-1]                                    \\
                \extradepscohs[n][p-1,...,n]:\extradepscohstype[n][p-1,...,n](\depscohs[n][0,...,p-1]) \\

@@

              \depscohs[n][0,...,p]:\depscohstype[n][0,...,p]                                        \\
              \extradepscohs[n][p-1,...,n]:\extradepscohstype[n][p-1,...,n](\depscohs[n][0,...,p-1]) \\

>>>>

\newcommandx{\deps}[2][1,2]{\prim{deps}[][#1][#2][][]}
\newcommandx{\depscohs}[2][1,2]{\prim{depscohs}[][#1][#2][][]}
\newcommandx{\depstype}[2][1,2]{\prim{depstype}[][#1][#2][][]}
\newcommandx{\depscohstype}[2][1,2]{\prim{depscohstype}[][#1][#2][][]}

@@

Formally, this can be described as:
\begin{equation*}
  \fontsize{7.2}{9}\selectfont
  \begin{array}{llcl}
    \framep[n][0,...,p]           &
    \multicolumn{3}{l}{
      \left(\begin{array}{l}
                \framep[n][0,...,p-1]:\frametype[n][0,...,p-1]                             \\
                \painting[n][0,...,p-1]:\paintingtype[n][0,...,p-1](\framep[n][0,...,p-1]) \\
                \restrf[n][0,...,p-1]:\restrftype[n][0,...,p-1]
                \left(\begin{array}{l}
                    \framep[n][0,...,p-1]   \\
                    \painting[n][0,...,p-1] \\
                  \end{array}\right)                                              \\
              \end{array}\right) : \frametype[n][0,...,p]
    }                                                                  \\
    \framep[n][0]                 &
    \left(\begin{array}{l}
              \unitpoint \\
              \unitpoint \\
              \unitpoint \\
            \end{array}\right)        & \defeq & (\unitpoint,unit)       \\
    \framep[n][0,...,p+1]         &
    \left(\begin{array}{l}
              \framep[n][0,...,p]   \\
              \painting[n][0,...,p] \\
              \restrf[n][0,...,p]   \\
            \end{array}\right)      & \defeq &
    \left(\begin{array}{l}
              \framep[n+1][0,...,p]
              \left(\begin{array}{l}
                  \framep[n+1][0,...,p-1]   \\
                  \painting[n+1][0,...,p-1] \\
                  \restrf[n+1][0,...,p-1]   \\
                \end{array}\right), \\
              \Sigma d:\framep[n+1][0,...,p]
              \left(\begin{array}{l}
                  \framep[n+1][0,...,p-1]   \\
                  \painting[n+1][0,...,p-1] \\
                  \restrf[n+1][0,...,p-1]   \\
                \end{array}\right).\, \\
              \quad \layer[n][p]
              \left(\begin{array}{l}
                  \framep[n+1][0,...,p-1]   \\
                  \painting[n+1][0,...,p-1] \\
                  \restrf[n+1][0,...,p-1]   \\
                \end{array}\right) \\
            \end{array}\right)                   \\
    \restrftype[n][0,...,p-1]     &
    \multicolumn{3}{l}{
      \left(\begin{array}{l}
                \framep[n][0,...,p-1]:\frametype[n][0,...,p-1]                             \\
                \painting[n][0,...,p-1]:\paintingtype[n][0,...,p-1](\framep[n][0,...,p-1]) \\
              \end{array}\right) : Type_{l}
    }                                                                  \\
    \restrftype[n][0,...,0-1]     &
    \left(\begin{array}{l}
              \unitpoint \\
              \unitpoint \\
            \end{array}\right)        & \defeq &
    \unittype                                                          \\
    \restrftype[n][0,...,(p+1)-1] &
    \left(\begin{array}{l}
              \framep[n][0,...,p]   \\
              \painting[n][0,...,p] \\
            \end{array}\right)      & \defeq &
    \begin{array}{l}
      \Sigma \restrf[n+1][0,...,p-1]
      :\restrftype[n+1][0,...,p-1]
      \left(\begin{array}{l}
                \framep[n+1][0,...,p-1]   \\
                \painting[n+1][0,...,p-1] \\
              \end{array}\right).                     \\
      \Pi q\leq n.\,\Pi \omega.\,\framep[n+1][p]
      \left(\begin{array}{l}
                \framep[n+1][0,...,p-1]   \\
                \painting[n+1][0,...,p-1] \\
                \restrf[n+1][0,...,p-1]   \\
              \end{array}\right) \rightarrow  \framep[n][p] \\
    \end{array} \\
    \layer[n][p]                  &
    \left(\begin{array}{l}
              \framep[n][0,...,p]     \\
              \painting[n][0,...,p]   \\
              \restrf[n+1][0,...,p-1] \\
            \end{array}\right)    & \defeq &
    \Pi\omega.\,\painting[n][p](\restrf[n][p](d))
  \end{array}
\end{equation*}

@@

This allows to define $\painting[n][p]$ formally:
\begin{equation*}
  \fontsize{7.2}{9}\selectfont
  \begin{array}{llcl}
    \painting[n][p]          &
    \multicolumn{3}{l}{
      \left(\begin{array}{l}
                \deps[n-1][0,...,n-1]:\depstype[n-1][0,...,n-1] \\
                E:\framep[n-1][n-1]\rightarrow Type_l           \\
              \end{array}\right) : \framep[n-1][n-1]\rightarrow Type_l
    }                                                            \\
    \painting[n][n]          &
    \left(\begin{array}{l}
              \deps[n-1][0,...,n-1] \\
              E                     \\
            \end{array}\right) & \defeq & E                        \\
    \painting[n][p](p<n)     &
    \left(\begin{array}{l}
              \deps[n-1][0,...,n-1] \\
              E                     \\
            \end{array}\right) & \defeq & d \mapsto
    \left(\begin{array}{l}
              \Sigma l:\layer(\deps[n-1][0,...,p-1]). \\
              \painting[n][p+1,...,n-1]
              \left(\begin{array}{l}
                  \deps[n-1][0,...,n-1] \\
                  E                     \\
                \end{array}\right)(d,l)           \\
            \end{array}\right) \\
  \end{array}
\end{equation*}

@@

                \deps[n][0,...,n]:\depstype[n][0,...,n] \\

@@

                    \deps[n][0,...,n] \\

@@

              \deps[n][0,...,n]   \\

@@

                \deps[n+1][0,...,n]     \\

@@

              \end{array}\right),                                             \\

@@

                \deps[n][0,...,n]:\depstype[n][0,...,n]                              \\

@@

              \end{array}\right).                          \\

@@

              \deps[n][0,...,n]:\depstype[n][0,...,p] \\

@@

At this new stage, it is convenient to use an abbreviation to combine sequences $\deps[n][0,..,n]$, $\restrp[n][0,...,p-1]$ and $\cohf[n][0,...,p-1]$ in one block, which we call $\depscohs[n][0,...p-1]$:

\begin{align*}
  \depscohstype[n][0,..,p-1] \defeq
  \left(\begin{array}{l}
            \deps[n][0,...,n] :\depstype[n][0,...,n]        \\
            \restrp[n][0,...p-1]: \restrptype[n][0,...,p-1] \\
            \cohf[n][0,...,p-1]: \cohftype[n][0,...,p-1]    \\
          \end{array}\right)
\end{align*}
so that the type of $\restrf[n][0,...,p]$ can be rephrased as $\depscohs[n][0,..,p-1] \rightarrow \restrftype[n][0,...,p]$. Later, we shall write $\depscohs.\deps$, $\depscohs.\restrp$, and $\depscohs.\cohf$ for the corresponding projections of a $\depscohs$.

@@

                \depscohs[n][0,...,n]:\depscohstype[n][0,...,n] \\

@@

                    \depscohs[n][0,...,n] \\

@@

              \depscohs[n][0,...,n] \\

@@

                \depscohs[n+1][0,...,n] \\

@@

                \depscohs[n][0,...,n]:\depscohstype[n][0,...,n]                          \\

@@

              \depscohs[n][0,...,n]:\depscohstype[n][0,...,n] \\
